---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# contourPolys

The goal of contourPolys is to create polygons via contourLines. 

Currently this is just me experimenting with the problem and keeping notes. 

## Example


By hacking `filled.contour` we can get all the fragments out, and plot properly with ggplot2. 


The *very* primitive `fcontour` function will provide the equivalent of the
data set produced and plotted by filled.contour. 

First, try to use `stat_contour`, it doesn't work because `contourLines` is
not producing closed regions, and worse the polygon drawing is not respecting holes that are filled by other smaller polygons. 

```{r}
## from https://twitter.com/BrodieGaslam/status/988601419270971392
library(reshape2)
v <- volcano
vdat <- melt(v)
names(vdat) <- c("x", "y", "z")
library(ggplot2)
ggplot(vdat, aes(x, y, z = z)) + 
  stat_contour(geom = "polygon",  aes(fill = ..level..))


cl <- contourLines(volcano)
image(volcano, col = NA); purrr::walk(cl, polygon)
```

One way to fix that is to *seal* the contour lines at the edges of the grid, but it's not easy to do. (The contouring is awesome in R, but the coordinates don't exactly go to the edges, which is the part I couldn't see an easy fix for).

A cheat's way, is to use a version of `filled.contour` and save all the fragments explicitly, and then plot as a set of tiny polygons. 

```{r}
z <- as.matrix(volcano)
y <- seq_len(ncol(z))
x <- seq_len(nrow(z))

levels <- pretty(range(z), n = 7)
p <- contourPolys::fcontour(x, y, z, levels)
m <- cbind(x = unlist(p[[1]]), 
           y = unlist(p[[2]]), 
           lower = rep(unlist(p[[3]]), lengths(p[[1]])), 
           upper = rep(unlist(p[[4]]), lengths(p[[1]])), 
           g = rep(seq_along(p[[1]]), lengths(p[[1]]))) 

gd <- as.data.frame(m)

library(ggplot2)
system.time({
print(ggplot(gd, aes(x, y, group = g, fill  = upper)) + geom_polygon())
})



```



Gggplot2 does plot many tiny polygons reasonably efficiently, because `grid::grid.polygon` is vectorized for aesthetics and for holes - but at some point it just won't scale for very many pixels. Ultimately we will want the regions as bounded areas.



We can coalesce these into efficient sf polygons, this is not done in an efficient way but there are 
improvements that could be made. 

* return a different organization of the fragments
* possibly, convert to edge-form, and simply remove any internal edges, then trace the remnants around in polygons (but that still needs to re-nest holes which is a hassle)
* build per level in C, rather than return all the fragments to R as a set, so the tighter the levels the smaller the overall footprint at any time
* some better marching squares proper thing ...


(This does work, try it at home ...)


```{r}
z <- as.matrix(volcano)
y <- seq_len(ncol(z))
x <- seq_len(nrow(z))

levels <- pretty(range(z), n =10)
p <- contourPolys::fcontour(x, y, z, levels)
m <- cbind(x = unlist(p[[1]]), 
           y = unlist(p[[2]]), 
           lower = rep(unlist(p[[3]]), lengths(p[[1]])), 
           upper = rep(unlist(p[[4]]), lengths(p[[1]])), 
           g = rep(seq_along(p[[1]]), lengths(p[[1]]))) 

r1 <- function(x) {
                   nr <- length(x)/2; 
structure(list(matrix(x, ncol = 2)[c(seq_len(nr), 1), ]), 
                             class = c("XY", "POLYGON", "sfg"))
}
library(sf)
xx <- lapply(split(m[, 1:2], rep(m[, 5], 2)), r1)
## drop bad ones
uu <- unlist(lapply(xx, st_is_valid))

x <- lapply(split(xx[uu], m[!duplicated(m[,5]), 3][uu]), 
            sf::st_geometrycollection)


x <- st_sfc(x)
y <- st_sf(geometry = x, a = seq_along(x))
plot(st_union(y, by_feature = TRUE))
```


But, this way has other advantages because with all the fragments the coordinates can be reprojected in a way that various R image plotters cannot do. 
(R needs some intermediate between array structures and polygons, so that shared vertices stayed shared (indexed) until needed, and expansion is progressively done while plotting/building areas, or we drop internal edges cleverly and trace around remaining boundaries. )


```{r}
# z <- volcano
# 
# x <- 10*1:nrow(z)
# y <- 10*1:ncol(z)
# d <- raster(list(x = x, y = y, z = z))
# 
# levels <- pretty(range(volcano))

library(raadtools)
d <- readtopo("etopo2", xylim = extent(120, 150, -45, -30))[[1]]
x <- yFromRow(d)
y <- xFromCol(d)
z <- as.matrix(d)
levels <- pretty(range(z), n = 7)
p <- contourPolys::fcontour(x, y, z, levels)
m <- cbind(x = unlist(p[[1]]), 
           y = unlist(p[[2]]), 
           lower = rep(unlist(p[[3]]), lengths(p[[1]])), 
           upper = rep(unlist(p[[4]]), lengths(p[[1]])), 
           g = rep(seq_along(p[[1]]), lengths(p[[1]]))) 

gd <- as.data.frame(m)
gd[c("x", "y")] <- proj4::ptransform(as.matrix(gd[c("y", "x")]) * pi/180, 
                                     "+init=epsg:4326", 
                                     "+proj=lcc +lon_0=147 +lat_0=-42 +lat_1=-30 +lat_2=-60")
library(ggplot2)
system.time({
print(ggplot(gd, aes(x, y, group = g, fill  = upper)) + geom_polygon())
})

## timing is okayish  
system.time({
library(grid)
grid.newpage()
cols <- viridis::viridis(length(levels))[scales::rescale(unlist(lapply(split(gd$lower, gd$g), "[", 1)), to = c(1, length(levels)))]
plot(range(gd$x), range(gd$y))
vp <- gridBase::baseViewports()
grid::pushViewport(vp$inner, vp$figure, vp$plot)
grid::grid.polygon(gd$x, gd$y,
                   gd$g,
                   gp = grid::gpar(fill = cols, col = NA),
                   default.units = "native")
grid::popViewport()
})

```


## Other attempts

Can we coalesce by detecting boundaries? 

We need

* find unique coordinates, and map UID to instances 
* find unique segments within region, segments identical despite order
* group_by region, segment and remove any segments that occur twice
* join all remaining segments, and coerce to polygon

Almost works, removing repeated segments certainly works - but still we have to re-nest the rings which is hard. 

Ultimately I think straight-through with lists of sanitized fragments into the indexed cascaded union will be fastest. 

```{r}
library(dplyr)
z <- as.matrix(volcano)
y <- seq_len(ncol(z))
x <- seq_len(nrow(z))

asub <- TRUE
if (asub) {
xsub <- seq(1, length(x), length = 17)
ysub <- seq(1, length(y), length = 16)
z <- z[xsub, ysub]
y <- y[ysub]
x <- x[xsub]
}
levels <- pretty(range(z), n = if(asub) 4 else 7)
p <- contourPolys::fcontour(x, y, z, levels)
m <- cbind(x = unlist(p[[1]]), 
           y = unlist(p[[2]]), 
           lower = rep(unlist(p[[3]]), lengths(p[[1]])), 
           upper = rep(unlist(p[[4]]), lengths(p[[1]])), 
           g = rep(seq_along(p[[1]]), lengths(p[[1]]))) 

gd <- tibble::as_tibble(m)

gd <- gd %>% group_by(g) %>% slice(c(1:n(), 1)) %>% ungroup()

 
# * find unique coordinates, and map UID to instances 
# * find unique segments within region, segments identical despite order
# * group_by region, segment and remove any segments that occur twice
# * join all remaining segments, and coerce to polygon

## clean up obvious degenerates
# udata <- gd %>%  
#   transmute(x, y, path = g, region = paste(lower, upper, sep = "-")) %>% 
#   unjoin::unjoin(x, y, key_col = ".vx")
# gd <- udata$data %>% group_by(path)  %>% distinct(.vx) %>% mutate(n = n()) %>% filter(n > 2) %>% 
#   ungroup() %>% 
#   select(path) %>% 
#   distinct() %>% 
#   inner_join(gd, c("path" = "g")) %>% 
#   mutate(path = as.integer(factor(path)))

udata <-   gd %>% transmute(x, y, path = g, region = paste(lower, upper, sep = "-")) %>% 
  unjoin::unjoin(x, y, key_col = ".vx")

segs <- purrr::map_df(split(udata$data$.vx, udata$data$path)[unique(udata$data$path)], silicate:::path_to_segment, .id = "path") 
segs$region <- udata$data$region[match(as.integer(segs$path), udata$data$path)]

## re-order segments to be sorted
vertex0 <- pmin(segs$.vertex0, segs$.vertex1)
vertex1 <- pmax(segs$.vertex0, segs$.vertex1)
segs$.vertex0 <- vertex0
segs$.vertex1 <- vertex1

usegs <- segs %>% mutate(segid = paste(.vertex0, .vertex1, sep = "-")) %>% 
  group_by(region, segid) %>% mutate(n = n()) %>% filter(n < 2) %>% ungroup()

tab <- usegs %>% inner_join(udata$.vx, c(".vertex0" = ".vx")) %>% rename(x0= x, y0 = y) %>% inner_join(udata$.vx, c(".vertex1" = ".vx"))
ggplot(tab , aes(x = x0, y = y0, xend = x, yend = y, col = region)) + geom_segment() + guides(colour = FALSE)

ucoords <- as.matrix(udata$.vx[c("x", "y")])
a <- purrr::map_df(split(usegs, usegs$region), 
           function(region) {
             #tibble::tibble(geometry = sf::st_sfc(sf::st_multilinestring( purrr::map(purrr::transpose(region[c(".vertex0", ".vertex1")]), ~ucoords[unlist(.x), ]))))
             
             tibble::tibble(geometry = sf::st_sfc(sf::st_multilinestring( purrr::map(purrr::transpose(region[c(".vertex0", ".vertex1")]), 
                                ~as.matrix(tibble(.vx = unlist(.x)) %>% inner_join(udata$.vx, ".vx") %>% select(x, y))))))
           }, .id = "region")


plot(sf::st_as_sf(a))


```

This seems to work, but the nesting is v hard to get right. 

```{r}
library(raster)
library(dplyr)
p2seg <- function(x) cbind(head(seq_len(nrow(x)), -1), 
                           tail(seq_len(nrow(x)), -1))
sf_explode <- function(x) {
  d <- sf::st_coordinates(x) %>% 
    tibble::as_tibble()
  Ls <- grep("^L", names(d))
  paster <- function(...) paste(..., sep = "-")
  dl <- d[-Ls] %>% mutate(path = do.call(paster, d[Ls])) %>% 
    split(.$path) 
  ll <- purrr::map(dl, ~lapply(split(t(p2seg(.x)), rep(seq_len(nrow(.x)-1), each = 2L)), 
                               function(idx) sf::st_linestring(as.matrix(.x[idx, c("X", "Y")])))) 
  sf::st_sfc(unlist(ll, recursive = FALSE))
}
#' Contour polygons
#'
#' @param x Raster
#' @param ... arguments passed to `contourLines`
#'
#' @return sf polygons
#' @export
#'
#' @examples
contour_poly <- function(x, levels = NULL, ..., nlevels = 10) {
  minmax <- c(raster::cellStats(x, min), raster::cellStats(x, max))
  if (is.null(levels)) levels <- seq(minmax[1] - 1, minmax[2], length = nlevels)
  ex <- raster::extend(x, 1L, value = minmax[1] - 1)
  cl <- rasterToContour(x, ...)
}

r <- extend(raster(volcano), 1, value = min(volcano)- 1)
cl <- rasterToContour(r, levels = seq(min(volcano) - 0.5, max(volcano) - 10, by = 20))
x <- sf_explode(sf::st_as_sf(cl))


library(sf)
p <- st_polygonize(st_union(x))
a <- st_cast(p)
st_overlaps(a)
plot(a, col = viridis::viridis(length(a)))


library(anglr)


```


